\ifspanish

\question[25] % 

Considere el problema de decisión dado por las verosimilitudes:
\begin{align}
p_{X|H}(x|1) &= \frac{\pi}{2} \sin\left(\frac{\pi}{2}x\right), \qquad 0<x<1 \nonumber\\
p_{X|H}(x|0) &= \frac{\pi}{2} \cos\left(\frac{\pi}{2}x\right), \qquad 0<x<1\nonumber
\end{align}

\begin{parts}
\part Determine las regiones de decisión del decisor LRT de parámetro $\eta$:
$$\frac{p_{X|H}(x|1)}{p_{X|H}(x|0)}\dunodcero \eta.$$
\part Represente gráficamente, de forma aproximada, la ROC del decisor LRT.
\part Represente, sobre la ROC, el punto de operación del decisor ML.
\part Represente, sobre la ROC, el punto de operación del decisor minimax.
\part Represente, sobre la ROC, el punto de operación del decisor de Neyman Pearson con $\pfa \le 0.4$.
\end{parts}

\begin{solution}
\begin{parts}
\part $x \dunodcero \dfrac{2}{\pi} \arctan(\eta)$, 
\part La ROC es un arco de circunferencia de radio 1 y centrado en (1,0).
\part $(\pfa,\pdet)= \left(1-\dfrac{\sqrt 2}{2},\dfrac{\sqrt 2}{2}\right)$
\part $(\pfa,\pdet)= \left(1-\dfrac{\sqrt 2}{2},\dfrac{\sqrt 2}{2}\right)$
\part $(\pfa,\pdet)= \left(0.4,0.8\right)$
\end{parts}
\end{solution}

\else

\question Consider a binary decision problem with likelihoods:
\begin{align}
p_{X|H}(x|1) &= \frac{\pi}{2} \sin\left(\frac{\pi}{2}x\right), \qquad 0<x<1 \nonumber\\
p_{X|H}(x|0) &= \frac{\pi}{2} \cos\left(\frac{\pi}{2}x\right), \qquad 0<x<1\nonumber
\end{align}

\begin{parts}
\part Find the decision regions of an LRT decision maker with parameter $\eta$:
$$\frac{p_{X|H}(x|1)}{p_{X|H}(x|0)}\dunodcero \eta.$$
\part Provide an approximate plot of the ROC of the LRT classifier.
\part Indicate which point of the ROC corresponds to the operating point of the ML classifier.
\part Indicate which point of the ROC corresponds to the operating point of the minimax decision maker.
\part Indicate which point of the ROC corresponds to the operating point of Neyman Pearson detector with $\pfa \le 0.4$.
\end{parts}

\begin{solution}
\begin{parts}
\part $x \dunodcero \dfrac{2}{\pi} \arctan(\eta)$, 
\part The ROC is a circumference arch of radius 1, and centered in (1,0).
\part $(\pfa,\pmis)= \left(1-\dfrac{\sqrt 2}{2},\dfrac{\sqrt 2}{2}\right)$
\part $(\pfa,\pmis)= \left(1-\dfrac{\sqrt 2}{2},\dfrac{\sqrt 2}{2}\right)$
\part $(\pfa,\pmis)= \left(0.4,0.8\right)$
\end{parts}
\end{solution}

\fi