\ifspanish

\question[20] % JCS 

Considere el problema de decisión binaria dado por la observación $X\in[0,2]$ y verosimilitudes
\[
p_{X|H}(x|1) = \frac{1}{2} x
\]
\[
p_{X|H}(x|0) = \frac{3}{4}x(2-x)^2,
\]
siendo $P_H(1)= \dfrac{2}{5}$. 

\begin{parts}
\part Determine el decisor MAP.
\part Determine la probabilidad de pérdida del decisor MAP.
\part Suponga ahora que el mismo decisor que se ha obtenido en el apartado (a) se aplica a un escenario en el que la verosimilitud de $H=1$ es
\[
p'_{X|H}(x|1) = \frac{7}{8} p_{X|H}(x|1)
              + \frac{1}{16},
\]
mientras que la verosimilitud de $H=0$ sigue siendo la misma. Determine el incremento en la probabilidad de error que se produce como consecuencia de este cambio de escenario.
\end{parts}

\begin{solution}
\begin{parts}
\part $x \dunodcero \dfrac{4}{3}$
\part $\pmis = \dfrac{4}{9}$
\part Dado que la $\pfa$ no cambia y $\pmis' = \dfrac{17}{36}$, el incremento de la probabilidad de error es $P_H(1)(\pmis'-\pmis)=\dfrac{1}{90}$
\end{parts}
\end{solution}

\else

\question Consider the binary decision problem characterized by an observation $X\in[0,2]$ and likelihoods
\[
p_{X|H}(x|1) = \frac{1}{2} x
\]
\[
p_{X|H}(x|0) = \frac{3}{4}x(2-x)^2,
\]
with $P_H(1)= \dfrac{2}{5}$. 

\begin{parts}
\part Find the MAP classifier.
\part Obtain the probability of missing of the MAP classifier.
\part Assume now that the same decision maker that was designed in subsection (a) is applied to a scenario in which the likelihood of $H=1$ is
\[
p'_{X|H}(x|1) = \frac{7}{8} p_{X|H}(x|1)
              + \frac{1}{16},
\]
whereas the likelihood of $H=0$ remains unchanged. Obtain the increment in the probability of error that takes place as a consequence of this different scenario.
\end{parts}

\begin{solution}
\begin{parts}
\part $x \dunodcero \dfrac{4}{3}$
\part $\pmis = \dfrac{4}{9}$
\part Since $\pfa$ does not change and $\pmis' = \dfrac{17}{36}$, the increment of the probability of error is $P_H(1)(\pmis'-\pmis)=\dfrac{1}{90}$
\end{parts}
\end{solution}

\fi