\ifspanish

\question Se dispone de tres variables aleatorias independientes $S_1$, $S_2$ y $S_3$, con idéntica fdp a priori. A partir de una observación de $X = S_1 + S_2 + S_3$ se desea estimar $S_1$. 

\begin{parts}
\part Justifique brevemente por qué el estimador de mínimo error cuadrático de $S_1$ dado $x$ es $\displaystyle \hat{s}_1 = \frac{x}{3}$.
\part Calcule el sesgo de dicho estimador. ¿Es sesgado o insesgado?
\vspace{0.4cm}

\hspace{-0.8 cm} Asumiendo en los apartados siguientes que $S_1$, $S_2$ y $S_3$ tienen una fdp uniforme entre -1 y 1:

\part Calcule la fdp de $X$ dado $s_1$.
\part Calcule la varianza del estimador $\hat{S}_1$.
\end{parts}

\begin{solution}
\begin{parts}
\part Por la simetría del problema, todas las $\mathbb{E}\{S_i|x\}$ deben ser iguales y sumar $x$.
\part 0, insesgado.
\part $p_{X|S_1}(x|s_1) = 1/2- |x-s_1|/4$ para $-2 < x < 2$, 0 en otro caso.
\part $1/9$
\end{parts}
\end{solution}

\else

\question Consider three independent random variables, $S_1$, $S_2$ and $S_3$, with the same {\em a priori} pdf. We wish to estimate $S_1$ from a single observation of $X = S_1 + S_2 + S_3$. 

\begin{parts}
\part Justify briefly why the minimum mean square error estimator of $S_1$ given $x$ is $\displaystyle \hat{s}_1 = \frac{x}{3}$.
\part Find the bias of the estimator given in (a).  Is it biased or unbiased?
\vspace{0.4cm}

\hspace{-0.8 cm} Assuming in the following that $S_1$, $S_2$ and $S_3$ follow a uniform distribution between $-1$ and $1$:

\part Obtain the pdf of $X$ given $s_1$.
\part Find the variance of estimator $\hat{S}_1$.
\end{parts}

\begin{solution}
\begin{parts}
\part Using symmetry arguments, all $\mathbb{E}\{S_i|x\}$ should be the same and sum up to $x$.
\part The estimator is unbiased.
\part $p_{X|S_1}(x|s_1) = 1/2- |x-s_1|/4$ for $-2 < x < 2$, and $0$ otherwise.
\part $\text{Var}(S_1) = 1/9$
\end{parts}
\end{solution}

\fi