\ifspanish

\question Se desea estimar la media de una v.a. $X$ a partir de un conjunto de $K$ observaciones independientes $\{X^{(k)}\}_{k=1}^{K}$, para lo que se construye el siguiente estimador:
$$\hat M = \frac{a}{K} \sum_{k=1}^K X^{(k)},$$
siendo $a$ una constante a determinar.
\begin{parts}
 		\part Calcúlense el sesgo y la varianza del estimador en función del valor de $a$.
        \part ¿Para qué valor de $a$ se minimiza la varianza? ¿Existe algún valor de $a$ para el que el estimador sea insesgado?
        \part Obténgase el error cuadrático medio del estimador y el valor de $a$ que lo minimiza.
\end{parts}

\begin{solution}
\begin{parts}
 		\part ${\mathbb E\left\{m - \hat M\right\} = (1 - a)} m$; $\text{Var}\left\{\hat M \right\} = \displaystyle\frac{a^2 v}{K}$.
        \part La varianza se minimiza para $a= 0$ y el sesgo es nulo para $a=1$.
        \part $\mathbb E\left\{(\hat M - m)^2 \right\} = (a - 1)^2 m^2 +\displaystyle\frac{a^2 v}{K} $.  El valor de $a$ que lo minimiza es $a^* = \displaystyle\frac{m^2}{m^2 + v/K}$.
\end{parts}
\end{solution}

\else

\question Consider the estimation of a r.v. $X$ from a set of $K$ independent observations $\{X^{(k)}\}_{k=1}^{K}$, and also consider the following estimator:
$$\hat M = \frac{a}{K} \sum_{k=1}^K X^{(k)},$$
with $a$ a constant to determine.
\begin{parts}
 		\part Obtain the bias and the variance of the estimator as a function of $a$.
        \part Which value of $a$ minimizes the variance? Is there any value of $a$ that produces an unbiased estimator?
        \part Obtain the mean square error of the estimator, and find the value of $a$ that minimizes it.
\end{parts}

\begin{solution}
\begin{parts}
 		\part $\mathbb E\left\{\hat M - m\right\} = (a - 1) m$; $\text{Var}\left\{\hat M \right\} = \displaystyle\frac{a^2 v}{K}$.
        \part The variance is minimized by $a= 0$, whereas the bias is null for $a=1$.
        \part $\mathbb E\left\{(\hat M - m)^2 \right\} = (a - 1)^2 m^2 +\displaystyle\frac{a^2 v}{K} $.  Thus, the value of $a$ minimizing it is $a^* = \displaystyle\frac{m^2}{m^2 + v/K}$.
\end{parts}
\end{solution}

\fi